# Missing values


```{r setup, include=FALSE}
# this prevents package loading message from appearing in the rendered version of your problem set
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)
```


```{r}
library(openintro)
library(tidyverse)
library(ggplot2)
library(GGally)
library(ggalluvial)
library(dplyr)
library(tidyverse)
library(patchwork)
```

## DataSet
In our data set of risk factors of cervical cancer, the missing values were original represented by question marks "?". We will first replace those marks with NA and then extract the first 100 rows as our example for this frist draft.
```{r }
df <- read_csv("risk_factors_cervical_cancer.csv", show_col_types = FALSE)
df <- df %>% na_if('?')
ex <- head(df, 100)
ex <- as.data.frame(apply(ex, 2, as.numeric))
```

## Missing Values by Column
```{r}
colSums(is.na(ex)) %>%sort(decreasing = TRUE)
```

## Missing Value Plots
Here is a missing value plots with heat map of the first 100 data rows. 
```{r }
library(mi)
names(ex) = abbreviate(names(ex), minlength = 2)
datasetdf <- as.data.frame(ex)
x <- missing_data.frame(datasetdf)
image(x)
```
Here is a missing value plots by variables of the first 100 data rows. 
```{r}
tidyex <- ex %>% 
    rownames_to_column("id") %>% 
    gather(key, value, -id) %>% 
    mutate(missing = ifelse(is.na(value), "yes", "no"))

ggplot(tidyex, aes(x = key, y = fct_rev(fct_reorder(id,as.integer(id))),
                   fill = value)) +
  geom_tile(color = "white") + 
  scale_fill_gradient(low = "grey80", high = "red", na.value = "black") +
  scale_y_discrete(guide = guide_axis(n.dodge=2)) +
  ylab("") +
  xlab("column names") +
  theme(axis.text.x = element_text(angle = 90, size = 6), 
        axis.text.y = element_text(size = 4))
```

## Using Problem 2 missing value function
```{r}
missing_plots = function(sample_data, percents){
  missing_patterns <- data.frame(is.na(sample_data)) %>%
    group_by_all() %>%
    count(name = "count", sort = TRUE) %>%
    ungroup() %>%
    rownames_to_column("id")
  missing_patterns = missing_patterns %>%
    rowwise() %>%
    mutate(complete = ifelse(sum(c_across(
    names(missing_patterns)[! names(missing_patterns) %in% 
                              c("count", "id")])) == 0, 1.1, 1))
  
  complete_row = missing_patterns[missing_patterns$complete == 1.1, ]$id
  
  if (percents){
    missing_patterns = missing_patterns %>%
      mutate(count = count / nrow(sample_data) * 100)
  }

  # gather
  tidy_pattern = missing_patterns %>%
    select(-count) %>%
    gather(key, value, -c(id, complete))

  # missing values by rows
  missing_rows = colSums(is.na(sample_data)) %>%
    sort(decreasing = TRUE)
  miss_rows=data.frame(missing_rows)
  miss_rows = miss_rows %>%
    rownames_to_column("id")
  if (percents){
    miss_rows = miss_rows %>%
      mutate(missing_rows = missing_rows/nrow(sample_data)*100)
  }
  
  # p1 is a plot of missing values by rows
  p1 = ggplot(miss_rows,aes(x = reorder(id, -missing_rows), 
                            y = missing_rows)) + 
    geom_col(fill = 'light blue') + 
    xlab('') +
    theme(axis.text.x = element_text(angle = 90, size = 5))
  if (percents){
    p1 = p1 + ylab('% rows\n missing') + 
      scale_y_continuous(limits = c(0,100))
  }
  else{
    p1 = p1 + ylab('num rows\n missing')
  }
  
  # p3 is a plot of missing patterns by count
  p3 = ggplot(missing_patterns, aes(x = count, 
                                    y = fct_reorder(id,
                                        as.integer(id))))+
    geom_col(fill = 'steelblue3', aes(alpha = complete)) +
    ylab('') +
    scale_y_discrete(limits=rev)+
    scale_alpha(range = c(0.5, 1.0))+
    theme(legend.position = "none")
  if (percents){
    p3 = p3 + xlab('% rows') +
      scale_x_continuous(limits = c(0,100))
  }
  else{
    p3 = p3 + xlab('row count')
  }
  
  # p2 is a plot of missing patterns by columns
  p2 = ggplot(tidy_pattern, aes(x = fct_reorder(key, -value, sum), y = fct_rev(fct_reorder(id,as.integer(id))), fill = value)) +
    geom_tile(color = "white", aes(alpha = complete))  +
    scale_fill_manual(values=c("gray", "mediumpurple1")) +
    xlab('variable') +
    ylab('missing patterns')+
    scale_alpha(range = c(0.5, 1.0))+
    theme(legend.position = "none", 
          axis.text.x = element_text(angle = 90, size = 5))
  
  if (length(complete_row) != 0){
    li = levels(fct_reorder(tidy_pattern$key, -tidy_pattern$value, sum))
    p2 = p2 + annotate('text', x = li[as.integer(length(li)/2)], 
                       y = complete_row, label = 'complete cases')
  }
  
  layout <- "
  AAAA#
  BBBBC
  BBBBC
  "
  graphs = p1 + p2 + p3 +
    plot_layout(design = layout)
  return (graphs)
}
```


```{r}
missing_plots(ex, TRUE)
```

As shown by the graph, we can see that both columns of name STDs: Time since first diagnosis and STDs: Time since last diagnosis contains a high number of missing data. We might need to looking into these two columns and decide whether to keep them given this high volume of NAs. Other missing values takes up some portion of each column.

