# Results
```{r}
library(tidyverse)
library(ggplot2)
library(GGally)
library(ggalluvial)
library(dplyr)
library(tidyverse)
library(patchwork)
```

```{r }
df <- read_csv("risk_factors_cervical_cancer.csv", show_col_types = FALSE)
df <- df %>% na_if('?')
colnames(df) = c("Ag", "Nosp","Fsi","Nop","Sm","SmY","SmPY","HC","HCY",
                 "IUD","IUDY","STD","STDN","STDcon","STDcer","STDvag",
                 "STDvul","STDsyp","STDpel","STDgen","STDmol","STDaid",
                 "STDhiv","STDhip","STDhpv","STDND","STDfd","STDld",
                 "DXc","DXcin","DXhpv","DX","Hn","Sch","Ct","Bp")
```

## Age

Many cancer diseases relate with age, that older people have higher risks. Is this consistent for cervical cancer in this data set?

The chance of one's having cervical cancer does not seem have a obvious connection to one's age from the four tests. From the graphs, the age overlapping rate is quite high between the two groups. The test results came out with "0" has more outlines and those test results with "1" is more concentrated around the whole IQR, this might results from the fact that the 1/0 ratio in out data is very small. Since, every median line of "1" box looks close to the median line of "0" box or lie outside of the whole IQR of box "0", there is unlikely to have a difference between the age factor and cervical cancer. From the graphs, we can also observe that the median of "1" is slightly higher than "0", meaning that the median of the group was tested with cervical cancer has a slightly higher age than those who do not have the disease.

```{r, echo = FALSE}
g_ag_bp = ggplot(df, aes(as.factor(Bp), Ag)) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Biopsy")+
  ylab("Age") +
  ggtitle("Age among Biopsy\n positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_ag_ct = ggplot(df, aes(as.factor(Ct), Ag)) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Cytology")+
  ylab("Age") +
  ggtitle("Age among Cytology\n positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_ag_sch = ggplot(df, aes(as.factor(Sch), Ag)) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Schiller")+
  ylab("Age") +
  ggtitle("Age among Schiller\n positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_ag_hn = ggplot(df, aes(as.factor(Hn), Ag)) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Hinselmann")+
  ylab("Age") +
  ggtitle("Age among Hinselmann\n positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

layout <- "
  AABB
  CCDD
  "

g_ag_bp + g_ag_ct + g_ag_sch + g_ag_hn + 
  plot_layout(design = layout)
```


## Number of Sexual Partners

Does number of sexual partners relate with cervical cancer? How do the 4 different ways of disease diagnosis show differences?

From the graphs, the Biopsy results and the Schiller results are very similar. However in Cytology results, the "0" box lies outside the "1" box, indicating there is likely to be a difference between the two groups. In all 3 tests, the median line of the "1" box seems close to the median line of "0" box except the Cytology graph. Their results and indications are quite consistent with each other, implying very little differences between the two groups.

```{r, echo = FALSE}
g_nosp_bp = ggplot(df, aes(as.factor(Bp), as.integer(Nosp))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Biopsy")+
  ylab("Number of Sexual Partners") +
  ggtitle("Number of Sexual Partners among Biopsy\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_nosp_ct = ggplot(df, aes(as.factor(Ct), as.integer(Nosp))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Cytology")+
  ylab("Number of Sexual Partners") +
  ggtitle("Number of Sexual Partners among Cytology\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_nosp_sch = ggplot(df, aes(as.factor(Sch), as.integer(Nosp))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Schiller")+
  ylab("Number of Sexual Partners") +
  ggtitle("Number of Sexual Partners among Schiller\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_nosp_hn = ggplot(df, aes(as.factor(Hn), as.integer(Nosp))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Hinselmann")+
  ylab("Number of Sexual Partners") +
  ggtitle("Number of Sexual Partners among Hinselmann\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

layout <- "
  AABB
  CCDD
  "

g_nosp_bp + g_nosp_ct + g_nosp_sch + g_nosp_hn + 
  plot_layout(design = layout)
```

## First sexual intercourse (age)

Does first sexual intercourse time correlate with cervical cancer?

Demonstrated by the graphs, in all 4 tests, the the median line of the "1" box seems to overlap with the median line of "0" box. Additionally, the two boxes in each graph has very high overlapping area, and they are very similar. This indicates that there is very unlikely to be a difference between the two groups, namely, very little connection between the factor of age of first sex intercourse and the chance of getting cervical cancer.

```{r}
g_fsi_bp = ggplot(df, aes(as.factor(Bp), as.integer(Fsi))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Biopsy")+
  ylab("First Sexual Intercourse (age)") +
  ggtitle("First Sexual Intercourse (age) among Biopsy\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_fsi_ct = ggplot(df, aes(as.factor(Ct), as.integer(Fsi))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Cytology")+
  ylab("First Sexual Intercourse (age)") +
  ggtitle("First Sexual Intercourse (age) among Cytology\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_fsi_sch = ggplot(df, aes(as.factor(Sch), as.integer(Fsi))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Schiller")+
  ylab("First Sexual Intercourse (age)") +
  ggtitle("First Sexual Intercourse (age) among Schiller\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_fsi_hn = ggplot(df, aes(as.factor(Hn), as.integer(Fsi))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Hinselmann")+
  ylab("First Sexual Intercourse (age)") +
  ggtitle("First Sexual Intercourse (age) among Hinselmann\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

layout <- "
  AABB
  CCDD
  "

g_fsi_bp + g_fsi_ct + g_fsi_sch + g_fsi_hn + 
  plot_layout(design = layout)
```


## Num of pregnancies

Pregnancy and giving birth affects cervical tissue in many ways. Would number of pregnancy relate to cervical cancer chances?

The bar plots shown in the following four tests all present a right-skewed trend with fewer people having high number of pregnancies. Both groups have high concentration of 1 to 3 number of pregnancies. The distributions for all four graphs are similar, however, it is clearly that the sample size of the group "1" is much smaller than the group "0" with a magnitude difference. 
```{r, echo = FALSE}
g_nop_bp = ggplot(df, aes(as.integer(Nop))) + 
  geom_histogram(color = "grey50", 
                fill = "lightblue",binwidth = 2) + 
  facet_wrap(~Bp, scales = "free") + 
  ylab("Count") + 
  xlab('Number of Pregnancies')+
  ggtitle("Number of Pregnancies among\n 
          Biopsy positive and negative") +
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_nop_ct = ggplot(df, aes(as.integer(Nop))) + 
  geom_histogram(color = "grey50", 
                fill = "lightblue",binwidth = 2) + 
  facet_wrap(~Ct, scales = "free") + 
  ylab("Count") + 
  xlab('Number of Pregnancies')+
  ggtitle("Number of Pregnancies among\n 
          Cytology positive and negative") +
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_nop_sch = ggplot(df, aes(as.integer(Nop))) + 
  geom_histogram(color = "grey50", 
                fill = "lightblue",binwidth = 2) + 
  facet_wrap(~Sch, scales = "free") + 
  ylab("Count") + 
  xlab('Number of Pregnancies')+
  ggtitle("Number of Pregnancies among\n 
          Schiller positive and negative") +
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_nop_hn = ggplot(df, aes(as.integer(Nop))) + 
  geom_histogram(color = "grey50", 
                fill = "lightblue",binwidth = 2) + 
  facet_wrap(~Hn, scales = "free") + 
  ylab("Count") + 
  xlab('Number of Pregnancies')+
  ggtitle("Number of Pregnancies among\n 
          Hinselmann positive and negative") +
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

layout <- "
  AABB
  CCDD
  "

g_nop_bp + g_nop_ct + g_nop_sch + g_nop_hn + 
  plot_layout(design = layout)
```

## Smoking

Smoking is related to many diseases and is seen as one of the most common factors to severe diseases including lung cancer and hear diseases. Does smoking relates to Cervical Cancer as well?

In the following part, we present a mosaic plot between boolean variables of smoking and biopsy results. In Biopsy test, the number of people who were diagnosed with cervical cancer or group "1" is roughly 10 times of that of those who were not, or the group "0" data. People who do not smoke is much more than people who smoke in this survey. In the group "1", the portion of people who smoke is slightly higher than of the the groups "0", but both are around 10 percent of the entire group.

```{r, fig.height = 3, fig.width = 3, fig.align = "center"}
library(vcd)
df$Sm = as.integer(df$Sm)
mosaic(Sm~Bp, data = df,direction = c("v", "h"),
       labeling = labeling_border(set_varnames = c(Bp = "Biopsy", 
                                   Sm = "Smoke")))
```

We calculate total packs of cigarettes smoked by using the equation where smoke years times Smoke packs per year. Presented by the graphs, we can clearly see that the most of the people smoke 0 packs of cigarettes since the IQR area of all the graphs are located at origin. And the people who were not diagnoised with cervical cancer or the group "0" has more outlines than those who were or the group "1". There is not clear difference between the two groups, indicating there is little relation between the factor of number of total packs of cigarettes smoked and the chance of getting cervical cancer. 
```{r, echo = FALSE}
df$SmT = as.double(df$SmY) * as.double(df$SmPY)
g_sm_bp = ggplot(df, aes(as.factor(Bp), as.double(SmT))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Biopsy")+
  ylab("Smoking packs total") +
  ggtitle("Smoking packs total among Biopsy\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_sm_ct = ggplot(df, aes(as.factor(Ct), as.double(SmT))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Cytology")+
  ylab("Smoking packs total") +
  ggtitle("Smoking packs total among Cytology\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_sm_sch = ggplot(df, aes(as.factor(Sch), as.double(SmT))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Schiller")+
  ylab("Smoking packs total") +
  ggtitle("Smoking packs total among Schiller\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_sm_hn = ggplot(df, aes(as.factor(Hn), as.double(SmT))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Hinselmann")+
  ylab("Smoking packs total") +
  ggtitle("Smoking packs total among Hinselmann\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

layout <- "
  AABB
  CCDD
  "

g_sm_bp + g_sm_ct + g_sm_sch + g_sm_hn + 
  plot_layout(design = layout)
```

## Hormonal Contraceptives

Hormonal Contraceptives are birth control methods, considered as safe and reliable ways to prevent pregnancy for most people. Would Hormonal Contraceptives be a factor for cervical cancer?

In the following graph, we have a mosaic plot between boolean variables Hormonal Contraceptives in biopsy results. In Biopsy test, the number of people who were diagnosed with cervical cancer or group "1" is roughly 10 times of that of those who were not, or the group "0" data. In the group "1", the portion of people who smoke is slightly higher than the the groups "0", but both are around 60 percent of the entire group.
```{r, fig.height = 3, fig.width = 3, fig.align = "center"}
df$HC = as.integer(df$HC)
mosaic(HC~Bp, data = df,direction = c("v", "h"),
       labeling = labeling_border(set_varnames = c(Bp = "Biopsy", 
                                   HC = "Hormonal Contraceptives")))
```

We investigate further on number of years of using Hormonal Contraceptives and its relation with cervical cancer. Shown by the graphs, in theses tests, the the median line of the "1" box seems to overlap with or very close to the median line of "0" box, except for Cytology test having a smaller median in label "1" box. Additionally, the two boxes in each graph has very high overlapping area, and they are very similar. This indicates that there is very unlikely to be a difference between the two groups, namely, very little connection between the factor of years of using Hormonal Contraceptives and the chance of getting cervical cancer.

```{r, echo = FALSE}
g_hc_bp = ggplot(df, aes(as.factor(Bp), as.double(HCY))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Biopsy")+
  ylab("Hormonal Contraceptives Years") +
  ggtitle("Hormonal Contraceptives Years among Biopsy\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_hc_ct = ggplot(df, aes(as.factor(Ct), as.double(HCY))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Cytology")+
  ylab("Hormonal Contraceptives Years") +
  ggtitle("Hormonal Contraceptives Years among Cytology\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_hc_sch = ggplot(df, aes(as.factor(Sch), as.double(HCY))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Schiller")+
  ylab("Hormonal Contraceptives Years") +
  ggtitle("Hormonal Contraceptives Years among Schiller\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_hc_hn = ggplot(df, aes(as.factor(Hn), as.double(HCY))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Hinselmann")+
  ylab("Hormonal Contraceptives Years") +
  ggtitle("Hormonal Contraceptives Years among Hinselmann\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

layout <- "
  AABB
  CCDD
  "

g_hc_bp + g_hc_ct + g_hc_sch + g_hc_hn + 
  plot_layout(design = layout)
```

## IUD

An IUD is a tiny device that's put into one's uterus to prevent pregnancy. It's long-term, reversible, and one of the most effective birth control methods. However, it is an extremely painful process and would make periods become heavier, more painful and longer. Question here would be if IUD relates with cervical cancer.

In the following graph, we have a mosaic plot between boolean variables Hormonal Contraceptives in biopsy results. In Biopsy test, the number of people who were diagnosed with cervical cancer or group "1" is roughly 10 times of that of those who were not, or the group "0" data. In the group "1", the portion of people who had IUD test seems to be higher than the the groups "0", but both are around 10 to 20 percent of the entire group.

A mosaic plot between boolean variables IUD and biopsy results.

```{r, fig.height = 3, fig.width = 3, fig.align = "center"}
df$IUD = as.integer(df$IUD)
mosaic(IUD~Bp, data = df,direction = c("v", "h"),
       labeling = labeling_border(set_varnames = c(Bp = "Biopsy", 
                                   IUD = "IUD")))
```

We investigate further on number of years of using IUD and its relation with cervical cancer. Demonstrated by the graphs, we can conclude that the most of the people spend 0 years using IUD since the IQR area of all the graphs are located at origin. And the people who were not diagnosed with cervical cancer or the group "0" has more outlines than those who were or the group "1". There is not clear difference between the two groups, indicating there is little relation between the factor of number of years of using IUD and the chance of getting cervical cancer. 

```{r, echo = FALSE}
g_iud_bp = ggplot(df, aes(as.factor(Bp), as.double(IUDY))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Biopsy")+
  ylab("IUD Years") +
  ggtitle("IUD Years among Biopsy\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_iud_ct = ggplot(df, aes(as.factor(Ct), as.double(IUDY))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Cytology")+
  ylab("IUD Years") +
  ggtitle("IUD Years among Cytology\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_iud_sch = ggplot(df, aes(as.factor(Sch), as.double(IUDY))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Schiller")+
  ylab("IUD Years") +
  ggtitle("IUD Years among Schiller\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_iud_hn = ggplot(df, aes(as.factor(Hn), as.double(IUDY))) +
  geom_boxplot() + 
  coord_flip() +
  xlab("Hinselmann")+
  ylab("IUD Years") +
  ggtitle("IUD Years among Hinselmann\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

layout <- "
  AABB
  CCDD
  "

g_iud_bp + g_iud_ct + g_iud_sch + g_iud_hn + 
  plot_layout(design = layout)
```

## STDs

Sexually transmitted diseases (STDs), or sexually transmitted infections (STIs), are infections that are passed from one person to another through sexual contact. How does STDs relate to cervical cancer? Is it a possible factor?




```{r, fig.height = 3, fig.width = 3, fig.align = "center"}
df$STD = as.integer(df$STD)
mosaic(STD~Bp, data = df,direction = c("v", "h"),
       labeling = labeling_border(set_varnames = c(Bp = "Biopsy", 
                                   STD = "STD")))
```

In the following graphs, we check if STD number correlates with cervical cancer by presenting a mosaic plot between boolean variables STD and biopsy results. The bar plots shown in the following four tests all present a right-skewed trend with both groups have most people clustered at origin. The distributions for all four graphs are similar, however, it is clearly that the sample size of the group "1" is much smaller than the group "0" with a magnitude difference. Since the distributions of both groups are ver similar, this indicates little differences between, namely, there is little connection between the STD number and the chance of getting cervical cancer.

```{r}
g_std_bp = ggplot(df, aes(as.double(STDN))) +
  geom_bar(color = "grey50", fill = "lightblue")+
  facet_wrap(~Bp, scales = 'free')+
  xlab("STD Number")+
  ggtitle("STD Number among Biopsy\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))
g_std_ct = ggplot(df, aes(as.double(STDN))) +
  geom_bar(color = "grey50", fill = "lightblue")+
  facet_wrap(~Ct, scales = 'free')+
  xlab("STD Number")+
  ggtitle("STD Number among Cytology\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_std_sch = ggplot(df, aes(as.double(STDN))) +
  geom_bar(color = "grey50", fill = "lightblue")+
  facet_wrap(~Sch, scales = 'free')+
  xlab("STD Number")+
  ggtitle("STD Number among Schiller\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

g_std_hn = ggplot(df, aes(as.double(STDN))) +
  geom_bar(color = "grey50", fill = "lightblue")+
  facet_wrap(~Hn, scales = 'free')+
  xlab("STD Number")+
  ggtitle("STD Number among Hinselmann\n 
          positive and negative")+
  theme_grey(14)+
  theme(plot.title = element_text(size = 10), 
        axis.title = element_text(size = 10))

layout <- "
  AABB
  CCDD
  "

g_std_bp + g_std_ct + g_std_sch + g_std_hn + 
  plot_layout(design = layout)
```

We would then look into a collection of different STDs and their correlation with cervical cancers. 

TODO!!!!!!!!!!


